---
title: "Download Guide Environmental Data to Sea Urchin"
subtitle: "Suplementary analysis to Stock Assessment Sea Urchin 2024"
author: "Mauricio Mardones"
date:  "`r format(Sys.time(), '%d %B, %Y')`"
bibliography: env_seaurchin.bib
csl: apa.csl
link-citations: yes
linkcolor: blue
output:
  pdf_document:
    toc: true
---
\pagebreak

## CONTEXT

Copernicus Marine data are available to download through server access services, all with the possibility of being integrated into scripts.
 
One of the most used endpoint is the MOTU API, by integrating the Python motuclient command line in a complete and operative script, regardless of the code language.
 
Indeed, motuclient is a cross-platform package that enables to download data by running a Python script. It can be integrated as a simple command line of your system from the language of your choice. 

Another method is to use the `CopernicusMarine` package developed to facilitate the retrieval of data from the Copernicus Marine catalog.
 
Let's see in this article an example of script to run a download process and how to use the `CopernicusMarine` package in R [@Good2020].

This data is download to do correlatrion analysis to get anothor factor to add in standarization in CPUE for sea urchin in Sourthern Chile fishery.

\pagebreak
## SETUP

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
Set libraries and Install first MOTU client

```{r eval=FALSE}
system("pip install motuclient==1.8.4")
#install.packages("CopernicusMarine")
```

```{r }
library(CopernicusMarine)
library(here)
```

Set credentials in Copernicus Data. 

```{r echo=FALSE}
CopernicusMarine_uid = "mmardones"           #ID
CopernicusMarine_pwd = "Trekkero_1"
```

Set dir data sink
Data folder is
```{r}
dir1 <- here("Data")
```

Define geo-frame to get data

We set limiit from X y XI region in Southern Chile

```{r echo=T, eval=T}
zona <- c(-81,-58,-70,-28)
```
Data available in the main repo from [Copernicus](https://data.marine.copernicus.eu/products)


The OSTIA [@Good2020] global sea surface temperature reprocessed product provides daily gap-free maps of foundation sea surface temperature and ice concentration (referred to as an L4 product) at 0.05deg.x 0.05deg. horizontal grid resolution, using in-situ and satellite data. This product provides the foundation Sea Surface Temperature, which is the temperature free of diurnal variability.

Then, select source data to analysis and put that name in the next line;


One of the variables to be analyzed and correlated with the biological data of the southern sea urchin will be chlorophyll, since this resource is a nutrient filter from the water mass. While there isn't a direct relationship, it is a test that must be done.

## Datos de clorofila


1 .Clhorophila (1997-2022) Global Ocean Colour (Copernicus-GlobColour), Bio-Geo-Chemical, L4 (monthly and interpolated) from Satellite Observations 
Overview

The biogeochemical hindcast for global ocean is produced at Mercator-Ocean (Toulouse. France). It provides 3D biogeochemical fields for the time period 1993-2019 at 1/4 degree and on 75 vertical levels. It uses PISCES biogeochemical model (available on the NEMO modelling platform). No data assimilation in this product.

Latest NEMO version (v3.6_STABLE)
Forcings: GLORYS2V4-FREE ocean physics produced at Mercator-Ocean and ERA-Interim atmosphere produced at ECMWF at a daily frequency
Outputs: Daily (chlorophyll. nitrate. phosphate. silicate. dissolved oxygen. primary production) and monthly (chlorophyll. nitrate. phosphate. silicate. dissolved oxygen. primary production. iron. phytoplankton in carbon) 3D mean fields interpolated on a standard regular grid in NetCDF format. The simulation is performed once and for all.
Initial conditions: World Ocean Atlas 2013 for nitrate. phosphate. silicate and dissolved oxygen. GLODAPv2 for DIC and Alkalinity. and climatological model outputs for Iron and DOC
Quality/Accuracy/Calibration information: See the related Quality Information Document
DOI (product): https://doi.org/10.48670/moi-00019

Full name: Global ocean biogeochemistry hindcast
Product ID : GLOBAL_MULTIYEAR_BGC_001_029
DOI:10.48670/moi-00019
Source :Numerical models
Spatial extent : Global OceanLat -90° to 90°Lon -180° to 180°
Spatial resolution: 0.25° × 0.25°
Temporal extent: 1 Jan 1993 to 31 Dec 2020
Temporal resolution: Daily-Monthly
Elevation (depth) levels: 75
Processing level: Level 4
Variables: 
Cell thickness Mass concentration of chlorophyll a in sea water (CHL)
Model level number at sea floorMole concentration of dissolved iron in sea water (FE)
Mole concentration of dissolved molecular oxygen in sea water (O2)
Mole concentration of nitrate in sea water (NO3)
Mole concentration of phosphate in sea water (PO4)
Mole concentration of phytoplankton expressed as carbon in sea water (PHYC)
Mole concentration of silicate in sea water (SI)
Net primary production of biomass expressed as carbon per unit volume in sea water (PP)
Sea floor depth below geoidSea water ph reported on total scale (pH)
Surface partial pressure of carbon dioxide in sea water (spCO2)
Feature type: Grid
Blue markets: Conservation & biodiversityClimate & adaptationPolicy & governanceScience & innovationMarine food
Projection: ETRS89 (EPSG 4258)
Data assimilation: None
Update frequency: Weekly
Format: NetCDF-4
Originating centre: Mercator Océan International
filename: cmems_mod.nc

Primero inspeccionamos el archivo y sus atributos.

```{r}
list1<-copernicus_products_list(freeText = "GLOBAL_ANALYSIS_FORECAST_BIO_001_028")  
detail1<-copernicus_product_details("GLOBAL_ANALYSIS_FORECAST_BIO_001_028")
```

Details for selecting subvariables 

```{r}
list1$mainVariables
list1$numLayers
detail1$layers$`cmems_mod_glo_bgc_my_0.25_P1D-m/chl`$variableTitle

```

Download `Mass concentration of chloropyll a in sea water`  data;

```{r eval=FALSE}
# Set parameters to subset and run the download 
copernicus_download_motu(
  username = CopernicusMarine_uid,
  password = CopernicusMarine_pwd,
  destination = dir1,
  product = "GLOBAL_ANALYSIS_FORECAST_BIO_001_028" ,
  layer = "global-analysis-forecast-bio-001-028-monthly",
  variable = "Mass concentration of chloropyll a in sea water",
  output = "netcdf",
  region  = zona,
  timerange = c("2000-01-01 12:00:00","2023-03-30 12:00:00"),
  sub_variables = c("chl"),
  verticalrange = c(0, 100)
)
```

Primero inspeccionamos el archivo y sus atributos.

```{r}
list1<-copernicus_products_list(freeText = "GLOBAL_MULTIYEAR_BGC_001_029")  
detail1<-copernicus_product_details("GLOBAL_MULTIYEAR_BGC_001_029")
```

Details for selecting subvariables 

```{r}
list1$mainVariables
list1$numLayers
detail1$layers$`cmems_mod_glo_bgc_my_0.25_P1D-m/chl`$subsetVariableIds

```
Bajo los datos de clorifila de todo chile desde el 2010 al 2022

```{r eval=FALSE}
copernicus_download_motu(
  username = "mmardones",
  password = "Trekkero_1",
  destination = "~",
  product = "GLOBAL_MULTIYEAR_BGC_001_029" ,
  layer = "cmems_mod_glo_bgc_my_0.25_P1D-m",
  variable = "Mass concentration of chloropyll a in sea water",
  output = "netcdf",
  region = c(-81,-58,-70,-28),
  timerange = c("2010-01-01 12:00:00","2022-03-30 12:00:00"),
  sub_variables = c("chl"),
  verticalrange = c(0, 100)
)
```
\pagebreak
# REFERENCES

