---
title: "Correlation Environmental and fishery sea urchin data"
subtitle: "Suplementary analysis to Stock Assessment Sea Urchin 2024"
author: "Mauricio Mardones I"
date:  "`r format(Sys.time(), '%d %B, %Y')`"
format:
  html:
    theme: cosmo
    fontsize: 0.9em
    linestretch: 1.7
    html-math-method: katex
    self-contained: true
    embed-resources: true
    code-tools: true
    toc: true
    toc-float: true
    toc-depth: 4
    link-citations: yes
linkcolor: blue
bibliography: env_seaurchin.bib
csl: apa.csl
---

```{r setup, include=FALSE, message=F}
rm(list = ls())
options(bitmapType = "cairo") 
#XQuartz is a mess, put this in your onload to default to cairo instead (https://github.com/tidyverse/ggplot2/issues/2655)
# Lo mapas se hacen mas rapido
# solo para IOs
```


```{r setup2, include=FALSE, message=F}
knitr::opts_chunk$set(collapse = TRUE, 
                      comment = "  ", 
                      fig.align = 'center',
                      cache=FALSE,
                      warning = FALSE,
                      message = FALSE)
```


# ANTECEDENTES





Cargo librerías necesarias para el análisis exploratorio de los datos de las distintas bases de bitácora, tallas y biológico. 

```{r lib, message=F, echo= TRUE}
library(ncdf4)
library(raster)
library(sf)
library(leaflet)
library(lubridate)
library(rasterVis)
library(tidyverse)
library(rnaturalearth)
library(rnaturalearthdata)
library(data.table)
library(marmap)
```

@natural2023; @ncdf42021; @raster2022; @lubridate2011
```{r}
# A function for dotplots
multi_dotplot <- function(filename, Xvar, Yvar){
  filename %>%
    ggplot(aes(x = {{Xvar}})) +
    geom_point(aes(y = {{Yvar}}),
               alpha=0.4) +
    theme_bw() +
    coord_flip() +
    labs(x = "Order of Data")}
```
## Datos de clorofila


1 .Clhorophila (1997-2022) Global Ocean Colour (Copernicus-GlobColour), Bio-Geo-Chemical, L4 (monthly and interpolated) from Satellite Observations 
Overview

The biogeochemical hindcast for global ocean is produced at Mercator-Ocean (Toulouse. France). It provides 3D biogeochemical fields for the time period 1993-2019 at 1/4 degree and on 75 vertical levels. It uses PISCES biogeochemical model (available on the NEMO modelling platform). No data assimilation in this product.

Latest NEMO version (v3.6_STABLE)
Forcings: GLORYS2V4-FREE ocean physics produced at Mercator-Ocean and ERA-Interim atmosphere produced at ECMWF at a daily frequency
Outputs: Daily (chlorophyll. nitrate. phosphate. silicate. dissolved oxygen. primary production) and monthly (chlorophyll. nitrate. phosphate. silicate. dissolved oxygen. primary production. iron. phytoplankton in carbon) 3D mean fields interpolated on a standard regular grid in NetCDF format. The simulation is performed once and for all.
Initial conditions: World Ocean Atlas 2013 for nitrate. phosphate. silicate and dissolved oxygen. GLODAPv2 for DIC and Alkalinity. and climatological model outputs for Iron and DOC
Quality/Accuracy/Calibration information: See the related Quality Information Document
DOI (product): https://doi.org/10.48670/moi-00019

Full name: Global ocean biogeochemistry hindcast
Product ID : GLOBAL_MULTIYEAR_BGC_001_029
DOI:10.48670/moi-00019
Source :Numerical models
Spatial extent : Global OceanLat -90° to 90°Lon -180° to 180°
Spatial resolution: 0.25° × 0.25°
Temporal extent: 1 Jan 1993 to 31 Dec 2020
Temporal resolution: Daily-Monthly
Elevation (depth) levels: 75
Processing level: Level 4
Variables: 
Cell thickness Mass concentration of chlorophyll a in sea water (CHL)
Model level number at sea floorMole concentration of dissolved iron in sea water (FE)
Mole concentration of dissolved molecular oxygen in sea water (O2)
Mole concentration of nitrate in sea water (NO3)
Mole concentration of phosphate in sea water (PO4)
Mole concentration of phytoplankton expressed as carbon in sea water (PHYC)
Mole concentration of silicate in sea water (SI)
Net primary production of biomass expressed as carbon per unit volume in sea water (PP)
Sea floor depth below geoidSea water ph reported on total scale (pH)
Surface partial pressure of carbon dioxide in sea water (spCO2)
Feature type: Grid
Blue markets: Conservation & biodiversityClimate & adaptationPolicy & governanceScience & innovationMarine food
Projection: ETRS89 (EPSG 4258)
Data assimilation: None
Update frequency: Weekly
Format: NetCDF-4
Originating centre: Mercator Océan International
filename: cmems_mod.nc

Primero inspeccionamos el archivo y sus atributos.

```{r}
ncfile <- "Data/global-analysis-forecast-bio-001-028-monthly_1693389410225.nc"

# Import NetCDF
nc <- nc_open(ncfile)

# Print information about the NetCDF file
print(nc)

# Obtener una lista de las variables en el archivo
attributes(nc$var)
attributes(nc$dim)
# Obtener la variable temporal
#unique(month)


```


```{r}
# import NetCDF with raster
chl_single <- raster(ncfile)
# print a summary of the raster
chl_single


# plot raster dataset
plot(chl_single)
```

```{r}
# import multi-band NetCDF file
chl_multi <- brick(ncfile)

# print a summary of the brick
chl_multi
```
```{r}
# plot brick dataset
levelplot(chl_multi)
```

```{r}
# calculate average and SD
chl_mean <- calc(chl_multi, fun = mean)
chl_sd <- calc(chl_multi, fun = sd)

#We can then explore the generated maps

# plot raster dataset
plot(chl_mean, main = "Average chl")
```

```{r}
plot(chl_sd, main = "Standard deviation chl")
```

```{r}
# convert raster to data.frame
chl_df <- as.data.frame(chl_mean, xy=TRUE, na.rm=TRUE)
```

Then, we will use the rnaturalearth package to download a global country map to use in the plot.

```{r}
# import countries layer from Natural Earth
countries <- ne_countries(scale = "medium", returnclass = "sf")
```

Finally, we use the ggplot() function from ggplot2 package to create a customized map. Note that each map component is added with a +. You can find further information about the ggplot2 package here.

```{r}
# plot
ggplot()+
  # add raster layer
  geom_raster(aes(x=x, y=y, fill=layer), data=chl_df) +
  # define color palette of raster layer
  scale_fill_distiller(palette = "Spectral", name = "chl (mgr)") + 
  # add countries layers
  geom_sf(fill=grey(0.9), color=grey(0.6), lwd = 0.2, data=countries) +
  # define spatial extent
  coord_sf(xlim = range(chl_df$x), ylim = range(chl_df$y), expand = F, ndiscr = 500) +
  # labels
  labs(title = "Mass concentration of chloropyll a in sea water(chl)",
       subtitle = "Annual average estimated from monthly products for 2020",
       x = "Longitude",
       y = "Latitude") +
  # theme
  theme_bw() 
```
## Datos de clorofila


1 .Clhorophila (1997-2022) Global Ocean Colour (Copernicus-GlobColour), Bio-Geo-Chemical, L4 (monthly and interpolated) from Satellite Observations 
Overview

The biogeochemical hindcast for global ocean is produced at Mercator-Ocean (Toulouse. France). It provides 3D biogeochemical fields for the time period 1993-2019 at 1/4 degree and on 75 vertical levels. It uses PISCES biogeochemical model (available on the NEMO modelling platform). No data assimilation in this product.

Latest NEMO version (v3.6_STABLE)
Forcings: GLORYS2V4-FREE ocean physics produced at Mercator-Ocean and ERA-Interim atmosphere produced at ECMWF at a daily frequency
Outputs: Daily (chlorophyll. nitrate. phosphate. silicate. dissolved oxygen. primary production) and monthly (chlorophyll. nitrate. phosphate. silicate. dissolved oxygen. primary production. iron. phytoplankton in carbon) 3D mean fields interpolated on a standard regular grid in NetCDF format. The simulation is performed once and for all.
Initial conditions: World Ocean Atlas 2013 for nitrate. phosphate. silicate and dissolved oxygen. GLODAPv2 for DIC and Alkalinity. and climatological model outputs for Iron and DOC
Quality/Accuracy/Calibration information: See the related Quality Information Document
DOI (product): https://doi.org/10.48670/moi-00019

Full name: Global ocean biogeochemistry hindcast
Product ID : GLOBAL_MULTIYEAR_BGC_001_029
DOI:10.48670/moi-00019
Source :Numerical models
Spatial extent : Global OceanLat -90° to 90°Lon -180° to 180°
Spatial resolution: 0.25° × 0.25°
Temporal extent: 1 Jan 1993 to 31 Dec 2020
Temporal resolution: Daily-Monthly
Elevation (depth) levels: 75
Processing level: Level 4
Variables: 
Cell thickness Mass concentration of chlorophyll a in sea water (CHL)
Model level number at sea floorMole concentration of dissolved iron in sea water (FE)
Mole concentration of dissolved molecular oxygen in sea water (O2)
Mole concentration of nitrate in sea water (NO3)
Mole concentration of phosphate in sea water (PO4)
Mole concentration of phytoplankton expressed as carbon in sea water (PHYC)
Mole concentration of silicate in sea water (SI)
Net primary production of biomass expressed as carbon per unit volume in sea water (PP)
Sea floor depth below geoidSea water ph reported on total scale (pH)
Surface partial pressure of carbon dioxide in sea water (spCO2)
Feature type: Grid
Blue markets: Conservation & biodiversityClimate & adaptationPolicy & governanceScience & innovationMarine food
Projection: ETRS89 (EPSG 4258)
Data assimilation: None
Update frequency: Weekly
Format: NetCDF-4
Originating centre: Mercator Océan International
filename: cmems_mod.nc


Ahora transformar los datos a `data.frame`

```{r  data chl2}
# Abrir el archivo .nc
raster_nc <- brick("Data/cmems_obs-oc_glo_bgc-plankton_my_l4-multi-4km_P1M_1693477867605.nc", varname = "CHL")
# Obtener la variable temporal
dates <- getZ(raster_nc)
years <-  as.integer(format(as.Date(dates), "%Y"))
month <-  as.integer(format(as.Date(dates), "%m"))
unique(years)

chla <- stackApply(raster_nc, years, fun=mean)
class(chla)

extension <- extent(chla)
proyeccion <- crs(chla)


# ahora extraigo los valores por cada punto 
stack_zona_chl <- raster::as.data.frame(chla, xy = TRUE)
stack_zona_chl <- na.omit(stack_zona_chl)
#str(stack_zona_chl)
```

Saco un plot del comportamiento de la variable en todos los años
```{r plot chla1}
spplot(chla, 
       colorkey = list(space = "bottom"), 
       scales = list(draw = TRUE),
      main = "Chorophyla a",
      fill='white',
      names.attr=c("2000",  "2001",  "2002",  "2003", 
                      "2004", "2005", "2006", "2007", "2008", "2009",
                      "2010","2011", "2012", "2013", "2014", "2015", 
                      "2016", "2017" , "2018",  "2019", "2020",
                   "2021", "2022"))

```

Ahora debo manipuñar la data para dejar formato largo table y luego extraer la serie de tiempo.


```{r manidat}
# Primero cambio nombre de las columnas
colnames(stack_zona_chl) <- c("lon", "lat",
                      "2000",  "2001",  "2002",  "2003", 
                      "2004", "2005", "2006", "2007", "2008", "2009",
                      "2010","2011", "2012", "2013", "2014", "2015", 
                      "2016", "2017" , "2018",  "2019", "2020",
                   "2021", "2022")
# ahora pivoteo los datos a lo largo
datchl <- stack_zona_chl %>%
  pivot_longer(cols = c( "2000",  "2001",  "2002",  "2003", 
                      "2004", "2005", "2006", "2007", "2008", "2009",
                      "2010","2011", "2012", "2013", "2014", "2015", 
                      "2016", "2017" , "2018",  "2019", "2020",
                   "2021", "2022"), 
               names_to = "ANO", values_to = "CHLa")
```

Compruebo la tendencia general de los datos a traves de la media de clorofila
```{r manidatchl}
daa=datchl %>% 
               group_by(lon,  lat) %>% 
               summarize(mean_var=mean(CHLa))

# un grafico simple
aschl <- ggplot(daa,
                aes(y=mean_var, 
                    x=as.double(ANO))) +
  geom_point(color="#386641")+
  stat_smooth(method = "loess",
              span = 0.3,
              col="#386641")+
  theme_light()+
  xlab("")
mchl1 <- ggplot() +
  geom_histogram(data = datchl, 
                 aes(CHLa),
                 bins = 100)+
  theme_bw()
ggarrange(mchl1,aschl)
```
Hago un mapa con los shapes nuevos

```{r}
chi <- raster::getData("GADM", country = "CHL", level = 0)
chi@data
#
e <- extent(-77,-72,-46,-41)
#e2 <- extent(-70,-35,-68,-60)
rc <- crop(chi, e)
proj4string(rc) <- CRS("+init=epsg:4326")
rc3 <- st_as_sf(rc) # par dejarlo en formato geom_sf

# Saco bathymetria
bat <- getNOAA.bathy(-77,-72,-46,-41, res = 1)
```


Preparo la grilla para los diversos indicadores.

```{r}
Grid2<- rc3 %>% #rc3 es el plot base original linea 487
  sf::st_make_grid(cellsize = c(0.2,0.125)) %>% 
  sf::st_cast("MULTIPOLYGON") %>%
  sf::st_sf() %>% # objeto en spatial feature
  dplyr::mutate(cellid = row_number()) 

# la pruebo

ggplot() +
  geom_contour(data = bat,
               aes(x=x, y=y, z=z),
               breaks=c(0, -10, -20, -50, -100, -200, -1000),
               linewidth=0.2,
               colour="grey")+
  geom_sf(data = df3, 
          color="grey",  
          fill=NA, 
          size=0.3) +
  geom_sf(data = rc3,  fill=NA)+
  theme_minimal()

```


```{r}
#transformar los datos en un sf object. Aqui ojo por q no cambié el nombre de las tallas =LONGITUD

df3 <- st_as_sf(datchl, coords = c(1, 2),  
                  crs = 4326) 

# join data to grid; make sure to join points to grid
# the first object drives the output geometry
result4 <- rc3 %>%
  st_intersection(df3) %>% 
  group_by(cellid, ANO) %>% 
  summarise(meanchl = mean(CHla))%>% 
  filter(! is.na(meanchl))
```

```{r }
cp1 <- ggplot() +
  #geom_sf(data = polimo, color="red",  fill=NA)+
  geom_contour(data = bat,
               aes(x=x, y=y, z=z),
               breaks=c(0, -10, -20, -50, -100, -200, -1000),
               linewidth=0.2,
               colour="grey")+
  geom_sf(data=result4, 
          aes(fill = cut(meanchla,
                         breaks = seq(0, 3, by = 50))),
          color=NA) +
  scale_fill_brewer(labels =  seq(0, 500, by = 50), # if you must
                    palette = "Purples",
                    name = "CPUE erizo") +
  geom_sf(data = rc3, color="black", 
          fill="white")+
  theme_bw()+
  #facet_wrap(.~ANO_ARR, ncol = 3)+
  coord_sf(crs = 4326)+
  labs(x = "Longitude", y = "Latitude")+
  ylim(-45.7, -41.3)+
  xlim(-75.2, -72.6)
cp1
```



# REFERENCIAS